\vspace{-15mm} \fontsize{24pt}{10pt}\selectfont \textbf{Traversing symmetric NAT with predictable port allocation function}

### 2013-12-15 paper.tex
vashek comments, algorithm successful run

% ------
% BibTex for bibliography
\usepackage[numbers]{natbib}
%\usepackage[square,sort&compress,authoryear]{natbib}

% ------
% Math
\usepackage{amsthm}
\usepackage{amsmath, mathtools}
\usepackage{fixltx2e}
% \usepackage{wrapfig}
\usepackage{tikz}
\usepackage{float}

@@

% ------
% Maketitle metadata
\title{\vspace{-15mm}%
	\fontsize{24pt}{10pt}\selectfont
	\textbf{Traversing Symmetric NAT with predictable port allocation function}
	}	
\author{%
	\large
	\textsc{Du\v{s}an Klinec} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
	\normalsize	Faculty of Informatics \\
	\normalsize	Masaryk University, Brno \\
	\normalsize	\href{mailto:xklinec@mail.muni.cz}{xklinec@mail.muni.cz}
	\and
	\textsc{Vashek Matyáš} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
	\normalsize	Faculty of Informatics \\
	\normalsize	Masaryk University, Brno \\
	\normalsize	\href{mailto:matyas@mail.muni.cz}{matyas@mail.muni.cz}
	\vspace{-5mm}
	}
\date{}

@@

In many applications is peer-to-peer (P2P) connection of crucial importance and the quality of a service depends on 
it's parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
between communicating parties is necessary. If P2P connection is not possible to be established relay 
servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
of another node in the communication path can have negative effects on a communication channel quality 
(increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
has to scale with increasing demands of the network, interesting point for attack, etc.
Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
establishing a direct connection between parties is of high importance.

@@

\section{Network Address Translation}
The notation we will be using for the rest of the paper follows.

@@

\paragraph{Full cone NAT.} ~\\
Allocation table $\mathbb{AT} \subseteq \{ \mathbb{IP}_{in} \times \mathbb{PORT} \} \times \mathbb{PORT}$ holds associations between internal 
hosts socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
Filtering rule is
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}_{in}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

\paragraph{Address restricted cone NAT.} ~\\
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex}\} \times \mathbb{PORT}$ 
holds associations between internal hosts socket, external host IP address and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
Filtering rule is
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

\paragraph{Port restricted cone NAT.} ~\\
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
holds associations between internal hosts socket, external host socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
Filtering rule is
\begin{align*}
& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

@@

\emph{Allocation function (AF)} says how is the port allocated for new connection chosen. Assume $\mathbb{AT}=\emptyset$, for simplicity. 
AF can be \emph{predictable}, e.g. \emph{incremental}. If previously allocated port was $\text{PORT\textsubscript{nat}}$
the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually. Another type of AF is \emph{random}.
If the chosen port is already taken, AF iterates until the free one is found.

@@

\emph{Port sensitive allocation function} creates new allocation if a destination address and a port differs from the existing
allocations from the same source\footnote{for each new connection is allocated new a NAT port.}:
\begin{align*}
& \exists s \in \mathbb{SP}, p \in \mathbb{PORT}: (s, p) \in \mathbb{AT} \Rightarrow \\
& \Rightarrow ((\forall s^{\prime} \in \mathbb{SP}, s^{\prime} \neq s , \forall p^{\prime} \in \mathbb{PORT}: \\
& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
\end{align*}

@@

\section{Existing techniques}
Many different solutions for traversing non-symmetric NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or Terredo\ignore{tunneling}~\citep{rfc4380}.

Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. But in general 
it is not widely deployed and it cannot be relied upon.

Most of the applications use Interactive Connectivity Establishment~(ICE)~\citep{rfc5245} nowadays, what is a very powerful toolbox for traversing 
various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN~\citep{rfc5766} for it's operation.

% \section{Related work}
Some interesting algorithms for traversing symmetric NAT were proposed by Y. Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
with port sensitive allocation function is not covered. It mentions that sending more packets won't increase probability of establishing the new connection.
Our motivation is to show that it is actually possible.

The \citep{rfc5128} maps current state of the art in NAT traversal. But does not cover our case.

Also the academic research was done in this field, for example by Wei~\emph{et.~al.} in \citep{wei}., but their algorithm is using another technique compared
to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is in majority of cases considered
as a privileged operation in operating system and sufficient privileges are required for this modification.

@@

\section{UDP hole punching with STUN}
All algorithms here make use of the same principle, \emph{UDP hole punching}. 
Internet hosts cannot send packet to the internal host directly. 
Communication has to be initiated from the inside\footnote{Ignoring 
pre-set port forwarding, DMZ, uPnP, etc\dots}, what is problematic if two nodes behind 
NAT wants to communicate.

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B} respectively,
A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
respectively. By default an internal host does not known it's external address and NAT type thus technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if I'll get it, etc...} 
is used to learn it. In case of incremental NAT it is also needed to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending to the other party and starting the algorithm is
denoted as an \emph{initialization phase} or a \emph{silent period}\footnote{Since the algorithm itself does nothing} of the NAT traversal algorithm.

@@

Moreover we assume incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
state\footnote{One new STUN request to determine current port number. Creates new NAT allocation.}
each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
\[
C_i = \begin{dcases*}
         0 & if $i=0$ \\
         C_{i-1} + 1 + X_i, \; X_i \sim Po(\lambda T) & otherwise 
        \end{dcases*}
\]
Rewritten as $C_i = i + \sum_{j=1}^{i}X_i$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.

\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected in this time interval} is assumed to be 3~minutes, 
what is common value in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$} pretty fast.

@@

\begin{mydef}
Symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where \\
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is set of communicating parties
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm
\item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
parameters and additional information obtained in initial phase). 
In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ is length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively
\end{compactitem}
\end{mydef}
If the third parameter is not important for the algorithm it can be ignored in the notation.

@@

\paragraph{Intuition.} 
If we have $A(0,1) = (2,3)$ it means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
of the party $1$.

@@

\paragraph{Party change.} 
We define unary operation on communicating parties. 
\[
a \in \mathcal{U},  \neg a = \begin{dcases*}
         0 & when $a=1$\\
         1 & when $a=0$
        \end{dcases*}
\]

\subsection{Algorithm I - baby-step, giant-step}
We propose the algorithm with high success rate for small network loads, i.e.,
$\lambda \leq 0.03, T=10$~ms. When error becomes larger,
algorithm step is not fast enough to compensate errors and thus fails for larger $\lambda$.

@@

\subsection{Algorithm II - fixed destination port}
Algorithm with fixed destination port described in by \citep{Wang:2006:RSN:1156422.1156550}
changes the source port in order to create new holes in NAT in each step.

@@

\subsection{Algorithm III - expected port value}
Some more effective algorithms can be designed if the random process the network follows is known (and their
parameters).

@@

\subsection{Algorithm IV - Poisson sampling}
Algorithm with the best success rate in our simulation is based on sampling Poisson distribution.
The main idea is to let the algorithm simulate the Poisson process that the network of the
other party follows.

\begin{mydef}
Poisson sampling algorithm 
\begin{align*}
& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + \text{P}(\lambda T (1+s C(\lambda T)))
\end{align*} where $\text{P}(\mu) = X \sim Po(\mu)$, samples Poisson distribution with
parameter $\mu$, coefficient function $C(\lambda T)$ gives a constant 
coefficient\footnote{This function was found empirically by maximizing the success rate}.
Approximations of the $C(\lambda T)$ by an inverse logarithm and a quartic model:
\begin{align*}
C(x) = & ({0.163321 \cdot  ln(64.2568 \cdot x)})^{-1}\\
C(x) = & (0.172876+1.28162\cdot x-1.41256\cdot x^2 \\
       & +0.825093\cdot x^3-0.184726\cdot x^4)^{-1}
\end{align*}
\end{mydef}
% 
\subsection{Silent period in algorithms.} If the length of the silent period is too big
the algorithm can have problems with coping with a gap that occurred in NAT ports. Thus 
Algorithms III and IV estimates the starting offset caused by the silent period as an
expected value from $E[Po(\lambda \cdot t_1)] = \lambda\cdot t_1$.

@@

\paragraph{Algorithm I and II comparison.} 
Algorithm I has following benefits over II:
\begin{compactitem}
 \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~s on average.
 \item Can be stopped as soon as connection is established, overhead is proportional to~$\lambda$.
 \item Single source port is used, only one listening thread is needed to implement
it\footnote{NAT is abused to multiplex multiple connections to one source port} compared to the Alg. II.
Due to this fact a practical implementation is simple and lightweight w.r.t. needed system resources.
 \item Is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
\end{compactitem}

@@

>>>>

% ------
% BibTex for bibliography
\usepackage[square,numbers]{natbib}
%\usepackage[square,sort&compress,authoryear]{natbib}

% ------
% Math
\usepackage{amsthm}
\usepackage{amsmath, mathtools}
\usepackage{fixltx2e}
% \usepackage{wrapfig}
\usepackage{tikz}
\usepackage{float}
\usepackage{multirow}

\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption

@@

% ------
% Maketitle metadata
\title{\vspace{-15mm}%
	\fontsize{24pt}{10pt}\selectfont
	\textbf{Traversing symmetric NAT with predictable port allocation function}
	}	
\author{%
	\large
	\textsc{Du\v{s}an Klinec} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
	\normalsize	Faculty of Informatics \\
	\normalsize	Masaryk University, Brno \\
	\normalsize	\href{mailto:xklinec@fi.muni.cz}{xklinec@fi.muni.cz}
	\and
	\textsc{Vashek Matyáš} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
	\normalsize	Faculty of Informatics \\
	\normalsize	Masaryk University, Brno \\
	\normalsize	\href{mailto:matyas@fi.muni.cz}{matyas@fi.muni.cz}
	\vspace{-5mm}
	}
\date{}

@@

In many applications is peer-to-peer (P2P) connection of crucial importance and the quality of a service depends on 
it's parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
between communicating parties is necessary. If P2P connection is not possible to be established, relay 
servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
of another node in the communication path can have negative effects on a communication channel quality 
(increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
has to scale with increasing demands of the network, interesting point for attack, etc.
Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
establishing a direct connection between parties is of high importance.

@@

\section{Existing techniques}
Many different solutions for traversing non-symmetric NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or 
Interactive Connectivity Establishment~(ICE)~\citep{rfc5245}. %Terredo\ignore{tunneling}~\citep{rfc4380}.

Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. But in general 
it is not widely deployed and it cannot be relied upon in a general case.

Most of the applications use ICE\ignore{~\citep{rfc5245}} nowadays, what is a very powerful toolbox for traversing 
various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for it's operation.

% \section{Related work}
Some interesting algorithms for traversing symmetric NAT were proposed by Y. Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
with port sensitive allocation function is not covered. It mentions that sending more packets does not increase a probability of establishing a new connection.
Our motivation is to show that it is actually possible.

RFC 5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
The research by Wei~\emph{et.~al.} in \citep{wei}., is devoted to symmetric NAT traversal but their algorithm is using another technique compared
to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is usually considered
as a privileged operation in the operating system and super-user privileges are required for this modification.

\section{Network Address Translation}
The notation used for the rest of the paper follows.

@@

\paragraph{Full cone NAT.} ~\\
Allocation table $\mathbb{AT} \subseteq \{ \mathbb{IP}_{in} \times \mathbb{PORT} \} \times \mathbb{PORT}$ holds associations between internal 
hosts socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
Filtering rule is:
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}_{in}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

\paragraph{Address restricted cone NAT.} ~\\
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex}\} \times \mathbb{PORT}$ 
holds associations between internal hosts socket, external host IP address and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
Filtering rule is:
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

\paragraph{Port restricted cone NAT.} ~\\
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
holds associations between internal hosts socket, external host socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
Filtering rule is:
\begin{align*}
& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

@@

\emph{Allocation function (AF)} defines port allocation rules for new connections (not in $\mathbb{AT}$). Assume $\mathbb{AT}=\emptyset$, for simplicity. 
AF can be either \emph{predictable} or \emph{random}. In our setting predictable AF is assumed, e.g., \emph{incremental}. 
If previously allocated port was $\text{PORT\textsubscript{nat}}$
the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually. Another type of AF is \emph{random}.
If the chosen port is already taken, AF iterates until the free one is found.

@@

\emph{Port sensitive allocation function} creates new allocation if a destination address and a port differs from the existing
allocations from the same source:
\begin{align*}
& \exists s \in \mathbb{SP}, p \in \mathbb{PORT}: (s, p) \in \mathbb{AT} \Rightarrow \\
& \Rightarrow ((\forall s^{\prime} \in \mathbb{SP}, s^{\prime} \neq s , \forall p^{\prime} \in \mathbb{PORT}: \\
& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
\end{align*}

@@

@@

\section{UDP hole punching with STUN}
All algorithms here make use of the same principle, \emph{UDP hole punching}. 
Internet hosts cannot send packet to the internal host directly. 
Communication has to be initiated from the inside\footnote{Not taking
pre-set port forwarding, DMZ, uPnP, etc... into consideration. }, what is problematic if two nodes behind 
NAT wants to communicate.

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B} respectively,
A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
respectively. By default an internal host does not known it's external address and NAT type thus technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if I'll get it, etc...} 
is used to learn it. In case of incremental NAT it is also needed to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending to the other party and starting the algorithm is
denoted as an \emph{initialization phase} or a \emph{silent period}\ignore{\footnote{Since the algorithm itself does nothing.}} of the NAT traversal algorithm.

@@

Moreover we assume incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
state\footnote{One new STUN request to determine current port number, it creates a new NAT allocation record.}
each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
\[
C_i = \begin{dcases*}
         0 & if $i=0$ \\
         C_{i-1} + 1 + X_i, \; X_i \sim Po(\lambda T) & otherwise 
        \end{dcases*}
\]
Rewritten as $C_i = i + \sum_{j=1}^{i}X_j$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.

\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected in this time interval.} is assumed to be 3~minutes, 
what is common value in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} pretty fast.

@@

\begin{mydef}
Symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where \\
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is set of communicating parties.
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm.
\item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
parameters and additional information obtained in initial phase). 
In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ is length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively.
\end{compactitem}
\end{mydef}
If the third parameter is not important for the algorithm it can be ignored in the notation.

@@

\paragraph{Intuition.} 
Line $A(0,1) = (2,3)$ means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
of the party $1$.

@@

%\paragraph{Party change.} 
%We define unary operation on communicating parties. 
%\[
%a \in \mathcal{U},  \neg a = \begin{dcases*}
%         0 & when $a=1$\\
%         1 & when $a=0$
%        \end{dcases*}
%\]

\paragraph{Algorithm log \& successful run.} IP addresses of the hosts and NAT are known and fixed prior algorithm run thus omitted
from the notation for simplicity. For the following text is assumed maximum number of steps of the algorithm $s_{max}=1000$.

\begin{mydef}
\label{def:log}
Algorithm log $\mathcal{L}^A_{\mathcal{A}}$ is a set of tuples $(p_1, p_2, p_3) \in \mathcal{L}^A_{\mathcal{A}}$ produced
by NAT traversal algorithm $\mathcal{A}$ execution on host $A$, where: \\
%$\mathcal{A}(0, s, (\lambda, t_s)) = (p_1, p_2)$ 
%are generated by $\mathcal{A}$ for some parameters $\lambda, t_s$ in some step $s$. The $p_3$ is 
%an allocated port by $NAT^A$ for request $(p_1, p_2)$ in a given step. Algorithm log for host $B$ is defined analogically. 
%Furthermore from our setting the 
%following formula holds:\footnote{From definition of a port sensitive symmetric NAT.} %$\mathcal{L^{A}} \subseteq PORT_{in}^{A} \times PORT_{nat}^{B} \times PORT_{nat}^{A}$,

\begin{compactitem}
 \item $(p_1, p_2) = \mathcal{A}(0, s, (\lambda, t_s)) $ are generated by the algorithm $\mathcal{A}$ for some 
       parameters $(\lambda, t_s)$ in some step $s$ of the algorithm $\mathcal{A}$.
 \item $p_3$ is an allocated port by $NAT^A$ for request $(p_1, p_2)$ in a given step $s$. In our network model
       $p_3=C_s=C_{s-1} + 1 + X, X~\sim~Po(\lambda T)$.
 \item Algorithm log for host $B$ is defined analogically.
 \item Furthermore from our setting the following formula holds:\footnote{From definition of a port sensitive symmetric NAT.}
\end{compactitem} %
%
% \begin{compactitem}
% \item $p_1$ is a source port.
% \item $p_2$ is a destination port.
% \item $p_3$ is an allocated port by local NAT for this request.
% \item following formula holds\footnote{from definition of a port sensitive symmetric NAT}
% \end{compactitem}
%
\begin{align*}
& (p_1, p_2, p_3) \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
& p_3 = p_3^{\prime} \Leftrightarrow \left( p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
\end{align*}
\end{mydef}

\begin{mydef}
\label{def:match}
Algorithm $\mathcal{A}$ run is successful, i.e., established a direct connection between $A,B$ if\footnote{Derived from UDP hole punching.}:
\begin{align*}
& MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}}) \Leftrightarrow \exists(p_1, p_2, p_3) \in \mathcal{L}^{A}_{\mathcal{A}}: \\
& \exists(p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{B}_{\mathcal{A}}: p_2 = p_3^{\prime} \wedge p_2^{\prime} = p_3
%
%  \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
%& p_3 = p_3^{\prime} \left( \Leftrightarrow p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
\end{align*}
\end{mydef}
The task is to find an algorithm $\mathcal{A}$ such that the probability the $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
holds is maximized for minimal number of steps needed.

\subsection{Algorithm I -- baby-step, giant-step}
We propose the algorithm with high success rate for small network loads, i.e.,
$\lambda \leq 0.03, T=10$~ms. When error becomes larger,
algorithm step is not fast enough to compensate errors and thus fails for larger $\lambda$.

@@

\subsection{Algorithm II -- fixed destination port}
Algorithm with fixed destination port described in by \citep{Wang:2006:RSN:1156422.1156550}
changes the source port in order to create new holes in NAT in each step.

@@

\subsection{Algorithm III -- expected port value}
Some more effective algorithms can be designed if the random process the network follows is known (and their
parameters).

@@

\subsection{Algorithm IV -- Poisson sampling}
Algorithm with the best success rate in our simulation is based on sampling Poisson distribution.
The main idea is to let the algorithm simulate the Poisson process that the network of the
other party follows. Duplicates are removed since it does not lead to new port allocation.

\begin{mydef}
Poisson sampling algorithm

% \captionof{algorithm}{Euclid’s algorithm}\label{algo1}

\begin{algorithmic}[1]
  \algtext*{EndWhile}% Remove "end while" text
  \algtext*{EndIf}% Remove "end if" text
  \algtext*{EndFunction}% Remove "end if" text
  \newcommand{\LineIf}[2]{ \State \algorithmicif\ {#1}\ \algorithmicthen\ {#2}}% \algorithmicend\ \algorithmicif }
  \Function{init4}{$\lambda, T$} %\Comment{The g.c.d. of a and b}
    \State $B\gets [], s \gets 0$
    %\State $s\gets 0$
    \While{$len(B)< s_{max}$}
      \State $x\gets \text{P}(\lambda T (1+s C(\lambda T))$
      \State $s\gets s+1$
      \LineIf{$x \notin B$}{$B.append(x)$} %\If {$x \notin B$} \State $B.append(x)$ \EndIf
    \EndWhile
    \State \textbf{return} $B$
    %\Return $B$
   \EndFunction\\

   \State $M\gets init4(\lambda,T)$\\
   \State $u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:$
%     \State $\mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + M[s])$
   \Function{A}{$u, s, (\lambda, t_s)$} %\Comment{The g.c.d. of a and b}
    %\LineIf{$s = 0$}{$M = init4(\lambda,T)$} %\If {$x \notin B$} \State $B.append(x)$ \EndIf
    \State \textbf{return} $(0, \lambda t_s + M[s])$
   \EndFunction
\end{algorithmic}
%\begin{align*}
%& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
%& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + \text{P}(\lambda T (1+s C(\lambda T)))
%\end{align*} 
where $\text{P}(\mu) = X \sim Po(\mu)$, samples Poisson distribution with
parameter $\mu$, coefficient function $C(\lambda T)$ gives a constant 
coefficient\footnote{This function was found empirically by maximizing the success rate in simulations.}.
Approximations of the $C(\lambda T)$ by an inverse logarithm and a quartic model:
\begin{align*}
C(x) = & ({0.163321 \cdot  ln(64.2568 \cdot x)})^{-1}\\
C(x) = & (0.172876+1.28162\cdot x-1.41256\cdot x^2 \\
       & +0.825093\cdot x^3-0.184726\cdot x^4)^{-1}
\end{align*}
\end{mydef}
% 
\subsection{Silent period in algorithms.} If the length of the silent period is too big
the algorithm can have problems with coping with a gap that occurred in NAT ports. Thus 
Algorithms III and IV estimates the starting offset caused by the silent period as an
expected value $E[Po(\lambda \cdot t_1)] = \lambda\cdot t_1$.

@@

\paragraph{Algorithm I and II comparison.} 
Algorithm I has following benefits over II:
\begin{compactitem}
 \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~sec on average.
 \item Can be stopped as soon as connection is established, overhead is proportional to~$\lambda$.
 \item Single source port is used, only one listening thread is needed to implement
it\footnote{NAT is abused to multiplex multiple connections to one source port.} compared to the Alg. II.
Due to this fact a practical implementation is simple and lightweight w.r.t. needed system resources.
 \item Is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
\end{compactitem}

@@

\section{Real data modeling}
Data collected on university network by NetFlow probes were studied to test the correspondence to the Poisson process
w.r.t. newly created connections/allocations on NAT. NAT was simulated in our program using
NetFlow data to provide network traffic to the NAT. NetFlow data capture 1 hour of a network communication in two different
day hours (03 -- low peak, 13 -- high peak).

NAT state (newly opened connections in window $T=10, 50, 100~ms$)
was sampled multiple times in a row, Poisson and Negative Binomial distributions were fitted using Maximum Likelihood Estimation (MLE)
and goodness-of-fit test with Pearson $\chi^2$ test was performed with confidence level $\alpha=0.05$.
